Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-15 surfaces-write-up/surface-write-up.tex

Furthermore, we have relations

\todo{relations}
Canonical form of 
\[
	\prod_{\substack{0 \leq i \leq m + 1 \\ \vec{y}
	\in \mss_i}} (F_i^{\vec{y}})^{g_{\vec{y}}}
\]

\noindent
where the indices with
some nonzero $g_{\vec{y}}$ are consecutive.
\end{proof}

\todo{change thm to be inductive for adding pts, giving result on
effective divisors}
\todo{remark on not doing two points}

\ssec{Bounds for Arbitrary Divisors on Projective Space}

\begin{rem}
When $\langle f_1, \ldots, f_r \rangle = \langle x_0, \ldots, x_n \rangle$ (the irrelevant ideal), we may directly use the hypersurface method by considering $x_0, \ldots, x_n$ instead of the $f_i$'s. However, when $\sqrt{\langle f_1, \ldots, f_r \rangle} = \langle x_0, \ldots, x_n \rangle$ (or the saturation), we cannot obtain the same result from our modification of O'Dorney's method as seen in Example ~\ref{eg:radical}. Instead we must use ghost points and hypersurfaces.
\end{rem} \todo{Peter: What is the purpose of this remark?  The only method we have extends to ghost points so that this is satisfied.}

@@

\noindent
for all $d\in \mathbb{N}$ and $c_i \ge -\alpha_i d$ satisfying $\bida_j + \sum_{i = 1}
^n \bida_i c_i = 0$.

>>>>

We also have relations involving generators in consecutive
indices. Suppose $F_i^{\vec{v}}$ and $F_j^{\vec{w}}$ are
generators with $j = i$ or $j = i + 1$ and $\vec{v} \in
\mss_i, \vec{w} \in \mss_j$ with $\vec{v} \not\prec \vec{w}$.
Let $\vec{y}$ and $\vec{z}$ be the unique vectors in $\mss_i$ and
$\mss_j$, respectively, such that $\vec{y} + \vec{z} = \vec{v} +
\vec{w}$ and $\vec{y} \prec \vec{z}$ (i.e.~ the nonzero indices
of $\vec{y}$ followed by those of $\vec{z}$ give an increasing
sequence). Then we see that
\begin{align*}
	F_i^{\vec{v}} F_j^{\vec{w}} = x^{\vec{v} + \vec{w}}
																 \left(\frac{u^{d_i}}{x_k^{c_i}}\right)
																 \left(\frac{u^{d_j}}{x_k^{c_j}}\right)
															= x^{\vec{y} + \vec{z}}
																 \left(\frac{u^{d_i}}{x_k^{c_i}}\right)
																 \left(\frac{u^{d_j}}{x_k^{c_j}}\right)
															= F_i^{\vec{y}} F_j^{\vec{z}}
\end{align*}

\noindent 
which give the relations $L_{i, j}^{\vec{v}, \vec{w}}$.

Now, we may apply the relations $L_{i, j}^{\vec{v}, \vec{w}}$
to any monomial in the $F_i^{\vec{v}}$'s involving
indices that differ by at most $1$ to produce the canonical form
\begin{align*}
	(F_i^{\vec{v}_{(1)}})^{g_{\vec{v}_{(1)}}} \cdots
	(F_i^{\vec{v}_{(1)}})^{g_{\vec{v}_{(\ell_1)}}}
	(F_{i + 1}^{\vec{w}_{(1)}})^{g_{\vec{w}_{(1)}}} \cdots
	(F_{i + 1}^{\vec{w}_{(\ell_2)}})^{g_{\vec{w}_{(\ell_2)}}}
\end{align*}

\noindent
where $\vec{v}_{(1)} \prec \vec{v}_{(2)} \prec \ldots \prec
\vec{v}_{(\ell_1)} \prec \vec{w}_{(1)} \prec \ldots \prec
\vec{w}_{(\ell_2)}$. Consequently, the relations of form
$G_{i, j}^{\vec{v}, \vec{w}}$ and of form $L_{i, j}^{\vec{v},
\vec{w}}$ generate all the relations among the $F_{i}^{\vec{v},
\vec{w}}$.
\end{proof}

\todo{change thm to be inductive for adding pts, giving result on
effective divisors?}
\todo{remark on not doing two points}

\ssec{Bounds for Arbitrary Divisors on Projective Space}
\begin{rem}
When $\langle f_1, \ldots, f_r \rangle = \langle x_0, \ldots, x_n \rangle$ (the irrelevant ideal), we may directly use the hypersurface method by considering $x_0, \ldots, x_n$ instead of the $f_i$'s. However, when $\sqrt{\langle f_1, \ldots, f_r \rangle} = \langle x_0, \ldots, x_n \rangle$ (or the saturation), we cannot obtain the same result from our modification of O'Dorney's method as seen in Example ~\ref{eg:radical}. Instead we must use ghost points and hypersurfaces.
\end{rem} \todo{Peter: What is the purpose of this remark?  The only method we have extends to ghost points so that this is satisfied.}

@@

\noindent
for all $d \in \mathbb{N}$ and $c_i \ge -\alpha_i d$ satisfying $\bida_j + \sum_{i = 1}
^n \bida_i c_i = 0$.
